Considering scale in multilayer animal social networks

### 2020-06-11 paper/manuscript.Rmd
landcover is land cover

## Landscape data and habitat classification
landcover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]. Available landcover classification included nine habitat
types at 30 m resolution. We reclassified the landcover types into three
categories: foraging habitat, open habitat, and forest habitat using the R
package `raster` [@Hijmans_2019]. Foraging habitat consisted of lichen habitat,
open habitat consisted of wetland, rocky barrens, and anthropogenic habitat
types, while forest habitat consisted of coniferous forest, conifer scrub,
broad leaf forest, and mixed-wood forest. Water habitat was excluded from all
subsequent analyses.

@@

### Network types and construction
We generated a multilayer social network consisting `r DT[, uniqueN(get(idcol))]`
individual nodes and two aspects (season and landcover). Two season
elementary-layers (summer and winter) and three landcover elementary-layers
(forage, forest, and open) combined for a total of six layers. Single layers were
composed of individuals linked by social associations, for each season and landcover combination. Proximity-based social
network layers were generated using the R package `spatsoc` [@Robitaille_2019]
in `r version$version.string` [@R_Core_Team_2019].

@@

We used a series of metrics to characterize the multilayer networks, focusing on
the role of individuals, importance of individual layers, and the similarity of
different layers. To measure the role of individuals, we calculated degree
centrality and graph strength. Degree centrality is the number of direct
connections an individual has to other individuals in a network layer. Graph
strength is the degree weighted by the strength of association, in this case
SRI, in each layer-specific season and landcover context. Extending degree
centrality, we calculated multidegree, the sum of degree centrality of
individuals across layers, and neighbourhood, the sum of unique neighbours to
each individual across layers [@Berlingerio_2012; @Kivela_2014]. We evaluated
the relevance and redundancy of layers by combining the multidegree and
neighbourhood to calculate layer relevance and connective redundancy. Layer
relevance is the proportion of neighbours present in each layer compared to the
neighbourhood for each individual and connective redundancy is the proportion of
neighbours repeated in multiple layers, and [@Berlingerio_2012]. Finally, we
calculated the similarity of layers between seasons, across landcover classes
(e.g. summer open compared to winter open) using the Pearson correlation
coefficient [@Brodka_2018]. Network metrics were calculated using the R packages
`igraph` [@Csardi_2006], `asnipe` [@Farine_2019] and `data.table` [@Dowle_2019].
<!-- add used spatsoc -->

@@

To assess the influence of spatial scale on multilayer networks, we aggregated
our landcover raster at varying scales relevant to caribou ecology. We
aggregated the 30 m landcover raster using the R package `grainchanger`
[@Graham_2019] at five scales: 100 m, 250 m, 500 m, 750 m and 1000 m. Aggregation
occurred using a modal moving window method using a circular window
corresponding to the above scales [@Graham_2019a]. We assume 30 m and 100 m
represents fine-scale decision making for caribou during foraging, while
re-sampling at 500 m , 750 m and 1000 m represents the scale at which caribou tend to
select and avoid habitat (Bastille-Rousseau et al. 2017).

@@

Increasing landcover resolution resulted in a decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats. Individuals in summer layers had consistent weak connections
(Figure \@ref(fig:lcres)a). Mean graph strength increased consistently in winter
forest and open habitats but varied in winter forage varied across landcover
resolutions with an increase at 250 m and decrease at 500 m (Figure
\@ref(fig:lcres)a). Layer relevance was stable for all habitat classes in summer
layers; layer relevance was significantly higher in winter and stable in open
and forest habitats but decreased for forage habitats (Figure
\@ref(fig:lcres)b). <!--This pattern during the winter indicates that
individuals had fewer neighbours in forage habitat habitats compared to their
overall neighbourhood across habitats, likely due to the patchy nature of
forage-->Layer similarity was low but stable for forest and open, except for a
sharp decrease in layer similarity in open at 100 m and 250 m (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
summer and winter forage because, at all landcover resolutions greater than 30
m, there were no individuals connected in the network layers. Individuals still
used forage habitats, but given the lower availability due to increasing
landcover resolution, at a much lower rate. Connective redundancy was moderately
high but decreased slightly as landcover resolution increased, indicating that
many neighbours were the same between layers, but the proportion of neighbours
in each layer slightly decreased at increasing landcover resolutions (Figure
\@ref(fig:lcres)d). Overall, these results indicate the importance of matching
landcover resolution to scale of selection as well as ensuring the landcover
product used has a resolution fine enough to detect landscape features and
habitats relevant to study species, such as important patchy and rare habitats.

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics appeared to stabilize. Mean graph
strength was consistently weak in summer layers but varied in winter layers at
low number of observations, becoming stable at ~75 observations (Figure
\@ref(fig:nobs)a). Layer relevance in the summer was consistently low, but
winter layers increased between approximately 10  and 100  observations,
remaining stable with additional observations (Figure \@ref(fig:nobs)b). Like
coarse landcover resolution (Figure \@ref(fig:lcres)), low numbers of
observations were not sufficient for capturing individuals in all contexts and
therefore layer similarity could not be calculated below ~30 observations for
forest, approximately ~80 observations for open (Figure \@ref(fig:nobs)c). Layer
similarity was not possible to calculate for forage layers because there were no
connected individuals in summer forage. Once sufficient observations were
obtained, layer similarity showed some variability with increasing numbers of
observations for forest and was relatively stable in open layers. Connective
redundancy increased dramatically between ~10 and ~50 observations before
stabilizing with additional observations included (Figure \@ref(fig:nobs)d).
Given the observed trend of these multilayer metrics stabilizing after a certain
number of observations are included, the results suggest this sensitivity method
could be useful for determining the number of observations necessary for
sufficiently describing social dynamics across contexts.

@@

<!-- Figure: metrics by landcover resolution -->
(ref:lcres) Varying spatial scale in landcover resolution. For each landcover resolution, multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength of individuals in summer layers had consistent weak connections. Individuals in winter forest and open habitats increased consistently while winter forage varied with a peak at 250 m and valley at 500 m. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter forage layer relevance at higher landcover resolutions. C) Similarity of landcover network layers between seasons (e.g. open summer compared to open winter) forage and forest habitats but showed more variation in open habitats with increasing social thresholds. Note that layer similarity in forage is not shown because individuals did not interact at landcover resolutions greater than 30 m during the summer in forage habitats, so layer similarity could not be calculated. D) Mean connective redundancy across individuals showed a slight decreasing trend with increasing landcover resolution.

@@

<!-- Figure: Window length -->
(ref:winlen) Varying temporal scale in time window length. For each time window length (40 to 100 days in length increasing by 5 days in sequence), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals in summer layers was weak but stable across increasing lengths of time except for a slight increase in strength for open and forage habitats when the season lasted more than 70 days. Mean graph strength was relatively stable over increasing time window lengths for winter layers, though there was more variation in strength between landcover types as window length increased. B) Mean layer relevance increased for summer layers across habitat classes and was stable for winter layers. C) Similarity of layers across habitat classes and between seasons (e.g. open summer compared to open winter) was variable with increasing window lengths, particularly for forage and open habitat types. D) Mean connective redundancy across individuals increased marginally with increasing window length.

@@

Multilayer networks are complicated in that they can incorporate data as varied
as social interactions, landscape features, and seasons, all of which exist at
different scales and units of measurement. Our analysis of caribou multilayer
networks which incorporated high resolution GPS telemetry data and spatial landcover
data is grounded in the context of observed animal social networks, as the
scales we chose to explore, e.g., number of observations, sociality across space
and time, are ones that can be observed as has traditionally been done in
ethology. We highlight ways in which scale ought to be considered for both
observed and remotely-sensed data when building social networks. Furthermore,
GPS telemetry data provides a link between large scale landscape properties and
fine scale social interactions as animal movement occurs across the landscape,
and when taken into account with multiple individuals can indicate association
patterns. Thus, a layer incorporating relocation data could be a great boon
linking spatial and social layer in multilayer networks.

>>>>

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]<!--TODO: check ref-->. Available land cover classification included nine habitat
types at 30 m resolution. We reclassified the land cover types into three
categories: foraging habitat, open habitat, and forest habitat using the R
package `raster` [@Hijmans_2019]. Foraging habitat consisted of lichen habitat,
open habitat consisted of wetland, rocky barrens, and anthropogenic habitat
types, while forest habitat consisted of coniferous forest, conifer scrub,
broad leaf forest, and mixed-wood forest. Water habitat was excluded from all
subsequent analyses.

@@

### Network types and construction
We generated a multilayer social network consisting `r DT[, uniqueN(get(idcol))]`
individual nodes and two aspects (season and land cover). Two season
elementary-layers (summer and winter) and three land cover elementary-layers
(forage, forest, and open) combined for a total of six layers. Single layers were
composed of individuals linked by social associations, for each season and land cover combination. Proximity-based social
network layers were generated using the R package `spatsoc` [@Robitaille_2019]
in `r version$version.string` [@R_Core_Team_2019].

@@

We used a series of metrics to characterize the multilayer networks, focusing on
the role of individuals, importance of individual layers, and the similarity of
different layers. To measure the role of individuals, we calculated degree
centrality and graph strength. Degree centrality is the number of direct
connections an individual has to other individuals in a network layer. Graph
strength is the degree weighted by the strength of association, in this case
SRI, in each layer-specific season and land cover context. Extending degree
centrality, we calculated multidegree, the sum of degree centrality of
individuals across layers, and neighbourhood, the sum of unique neighbours to
each individual across layers [@Berlingerio_2012; @Kivela_2014]. We evaluated
the relevance and redundancy of layers by combining the multidegree and
neighbourhood to calculate layer relevance and connective redundancy. Layer
relevance is the proportion of neighbours present in each layer compared to the
neighbourhood for each individual and connective redundancy is the proportion of
neighbours repeated in multiple layers, and [@Berlingerio_2012]. Finally, we
calculated the similarity of layers between seasons, across land cover classes
(e.g. summer open compared to winter open) using the Pearson correlation
coefficient [@Brodka_2018]. Network metrics were calculated using the R packages
`igraph` [@Csardi_2006], `asnipe` [@Farine_2019] and `data.table` [@Dowle_2019].
<!-- add used spatsoc -->

@@

To assess the influence of spatial scale on multilayer networks, we aggregated
our land cover raster at varying scales relevant to caribou ecology. We
aggregated the 30 m land cover raster using the R package `grainchanger`
[@Graham_2019] at five scales: 100 m, 250 m, 500 m, 750 m and 1000 m. Aggregation
occurred using a modal moving window method using a circular window
corresponding to the above scales [@Graham_2019a]. We assume 30 m and 100 m
represents fine-scale decision making for caribou during foraging, while
re-sampling at 500 m , 750 m and 1000 m represents the scale at which caribou tend to
select and avoid habitat (Bastille-Rousseau et al. 2017).

@@

Increasing land cover resolution resulted in a decreased availability of forage,
a rare and patchy habitat, and corresponding low connectivity of individuals in
these habitats. Individuals in summer layers had consistent weak connections
(Figure \@ref(fig:lcres)a). Mean graph strength increased consistently in winter
forest and open habitats but varied in winter forage varied across land cover
resolutions with an increase at 250 m and decrease at 500 m (Figure
\@ref(fig:lcres)a). Layer relevance was stable for all habitat classes in summer
layers; layer relevance was significantly higher in winter and stable in open
and forest habitats but decreased for forage habitats (Figure
\@ref(fig:lcres)b). <!--This pattern during the winter indicates that
individuals had fewer neighbours in forage habitat habitats compared to their
overall neighbourhood across habitats, likely due to the patchy nature of
forage-->Layer similarity was low but stable for forest and open, except for a
sharp decrease in layer similarity in open at 100 m and 250 m (Figure
\@ref(fig:lcres)c). Layer similarity was not possible to calculate between
summer and winter forage because, at all land cover resolutions greater than 30
m, there were no individuals connected in the network layers. Individuals still
used forage habitats, but given the lower availability due to increasing
land cover resolution, at a much lower rate. Connective redundancy was moderately
high but decreased slightly as land cover resolution increased, indicating that
many neighbours were the same between layers, but the proportion of neighbours
in each layer slightly decreased at increasing land cover resolutions (Figure
\@ref(fig:lcres)d). Overall, these results indicate the importance of matching
land cover resolution to scale of selection as well as ensuring the land cover
product used has a resolution fine enough to detect landscape features and
habitats relevant to study species, such as important patchy and rare habitats.

As the number of observations used to generate multilayer networks increased,
individuals became more connected and metrics appeared to stabilize. Mean graph
strength was consistently weak in summer layers but varied in winter layers at
low number of observations, becoming stable at ~75 observations (Figure
\@ref(fig:nobs)a). Layer relevance in the summer was consistently low, but
winter layers increased between approximately 10  and 100  observations,
remaining stable with additional observations (Figure \@ref(fig:nobs)b). Like
coarse land cover resolution (Figure \@ref(fig:lcres)), low numbers of
observations were not sufficient for capturing individuals in all contexts and
therefore layer similarity could not be calculated below ~30 observations for
forest, approximately ~80 observations for open (Figure \@ref(fig:nobs)c). Layer
similarity was not possible to calculate for forage layers because there were no
connected individuals in summer forage. Once sufficient observations were
obtained, layer similarity showed some variability with increasing numbers of
observations for forest and was relatively stable in open layers. Connective
redundancy increased dramatically between ~10 and ~50 observations before
stabilizing with additional observations included (Figure \@ref(fig:nobs)d).
Given the observed trend of these multilayer metrics stabilizing after a certain
number of observations are included, the results suggest this sensitivity method
could be useful for determining the number of observations necessary for
sufficiently describing social dynamics across contexts.

@@

<!-- Figure: metrics by land cover resolution -->
(ref:lcres) Varying spatial scale in land cover resolution. For each land cover resolution, multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength of individuals in summer layers had consistent weak connections. Individuals in winter forest and open habitats increased consistently while winter forage varied with a peak at 250 m and valley at 500 m. B) Mean layer relevance showed sparse connectivity in summer compared to winter networks and a decrease in winter forage layer relevance at higher land cover resolutions. C) Similarity of land cover network layers between seasons (e.g. open summer compared to open winter) forage and forest habitats but showed more variation in open habitats with increasing social thresholds. Note that layer similarity in forage is not shown because individuals did not interact at land cover resolutions greater than 30 m during the summer in forage habitats, so layer similarity could not be calculated. D) Mean connective redundancy across individuals showed a slight decreasing trend with increasing land cover resolution.

@@

<!-- Figure: Window length -->
(ref:winlen) Varying temporal scale in time window length. For each time window length (40 to 100 days in length increasing by 5 days in sequence), multilayer networks were constructed of caribou social associations (*Rangifer tarandus*, n = `r DT[, uniqueN(get(idcol))]` females) on Fogo Island, Newfoundland in summer 2017 and winter 2018 across three habitat classes (forage, forest, and open). A) Mean graph strength across individuals in summer layers was weak but stable across increasing lengths of time except for a slight increase in strength for open and forage habitats when the season lasted more than 70 days. Mean graph strength was relatively stable over increasing time window lengths for winter layers, though there was more variation in strength between land cover types as window length increased. B) Mean layer relevance increased for summer layers across habitat classes and was stable for winter layers. C) Similarity of layers across habitat classes and between seasons (e.g. open summer compared to open winter) was variable with increasing window lengths, particularly for forage and open habitat types. D) Mean connective redundancy across individuals increased marginally with increasing window length.

@@

Multilayer networks are complicated in that they can incorporate data as varied
as social interactions, landscape features, and seasons, all of which exist at
different scales and units of measurement. Our analysis of caribou multilayer
networks which incorporated high resolution GPS telemetry data and spatial land cover
data is grounded in the context of observed animal social networks, as the
scales we chose to explore, e.g., number of observations, sociality across space
and time, are ones that can be observed as has traditionally been done in
ethology. We highlight ways in which scale ought to be considered for both
observed and remotely-sensed data when building social networks. Furthermore,
GPS telemetry data provides a link between large scale landscape properties and
fine scale social interactions as animal movement occurs across the landscape,
and when taken into account with multiple individuals can indicate association
patterns. Thus, a layer incorporating relocation data could be a great boon
linking spatial and social layer in multilayer networks.
